---
title: "Credit Scoring SBD2"
author: "Abishan Arumugavel"
date: "2023-10-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Used librarys
```{r}
library(readr)
```


## Import dataset
```{r show_col_types = False}
loan_sample <- read_csv("loan_sample_8.csv")
```


## Excercise 1

### Describe the data. Specifically:

– Check and report the structure of the data set.
– How many numeric and how many categorical variables are included in the data? What categorical variable has the most levels in it?
– Summarize the variables. Discuss the summary statistics obtained.
– Check the levels of the target variable by choosing the appropriate visualization. Is the target variable balanced?
– Check the distribution of the numeric variables in the data set (include different visual representations).



• Investigating outliers with a box plot
Elaborate your view on how to proceed in dealing with the outliers and – if necessary – take appropriate action.
```{r}


```

• Investigating the distribution of the numeric features per the two levels of our target feature
(i.e. default vs non-default). Discuss the visualizations. Which variables seem to be relevant in predicting the target feature?
```{r}


```

• Use a bar plot visualization to investigate the associations between the categorical variables and the target feature.
```{r}


```

• Visualize the correlations that emerge between the numerical features. Discuss the results. Which variables are highly correlated? Decide whether you keep all variables.
```{r}


```

• Plot an interactive scatter plot of the association between the loan amount requested and the annual income of the borrower. Discuss the plot. What can you tell about the association?
```{r}


```

• Create a new balanced data set where the two levels of the target variable will be equally represented; Create a bar plot of the newly created target variable. Why is this step necessary?
```{r}


```

